---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# delaydiscount

<!-- badges: start -->
<!-- badges: end -->

The goal of the delaydiscount package is to provide methods for analyzing delay discounting data using the linearized hyperbolic model.

## Installation

You can install the development version of delaydiscount from Github.

``` {r, message = FALSE}
# install.packages("devtools")
devtools::install_git("https://github.com/davideh-vt/delaydiscount")
```

## Example

To fit a model on a dataset of delay discounting data, first run `prepare_data_frame` on the dataset, then run `dd_hyperbolic_model` on the output.

```{r example}
library(delaydiscount)
## basic example of fitting a model
prep_remedi <- prepare_data_frame(remedi)
fit <- dd_hyperbolic_model(prep_remedi)
fit$ln_k_mean
fit$var
```
