---
title: "rule-check"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rule-check}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(delaydiscount)
```

TODO: This file should include the entire workflow and not just the rule check. However, I do not want to add it until I finish it for the main path (no rule check), since any edits will need to be copied over.

The jb_rule_check function can be used to apply the rule check for the purpose of filtering out subjects who fail.

```{r}
remedi_rc_result <- jb_rule_check(remedi)
head(remedi_rc_result)
```

The resulting data frame has an observation for each subject-within-group, with variables corresponding to whether or not each rule was passed. A value of TRUE indicates that the rule check was passed.

C1 fails if there are two or more increases in the indifference point of 20\% of the total amount between consecutive time points. C2 fails if the difference in the indifference point between the first and last time points is not more than 10\% of the total amount.

In order to make use of the rule check results, after the data frame is prepared, the rule check data can be merged with the original dataframe, and then observations corresponding to subjects who failed the rule check can be removed from the data.

```{r}
remedi_with_rc <- merge(remedi, remedi_rc_result)
remedi_rc_pass <- remedi_with_rc %>%
  dplyr::filter(C1 == TRUE, C2 == TRUE)
```

After that, the remainder of the workflow proceeds as 
