---
title: "simulate-dataset"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simulate-dataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(delaydiscount)
library(dplyr)
```

The `simulate_dataset` method can be used to simulate data from the hierarchical linearized hyperbolic model.

The hierarchical linearized hyperboloid model works as follows.

The hyperbolic model for the discounting curve is $D(t) = \frac{1}{1+kt}$. However, actual discounting rates reported by subjects may not fit this curve. Actual discounting proportions are assumed to be between 0 and 1, exclusive. They are transformed from the $(0,1)$ scale to a $(-\infty, \infty)$ scale by the function $\phi(x) = \log(\frac{1}{x}-1)$.

The model discounting curve with the transformation applied is $\log(\frac{1}{D(t)} - 1) = \log(k) + \log(t)$. The transformed curve provides a mean structure for the data, conditional on the subject's $\log(k)$ parameter.

We assume that the transformed observed indifference point $\tilde D(t_{ijc})$ will follow the model

$\log(\frac{1}{\tilde D(t_{ijc})} - 1) = \log(k_{ic}) + \log(t_{ijc}) + \epsilon_{ijc},$

where $\epsilon_{ijc} \sim N(0, \sigma^2)$, $i$ is an index for subject, $j$ an index for time point, and $c$ is an index for group/condition.

$\log(k_{ic}) \sim N(\theta_c, \frac{g\sigma^2}{T})$, where $\theta_c$ is the population mean of the $\log(k_{ic})$ for subjects in condition $c$, and $T$ is the number of time points observed for each subject.

The model is hierarchical, as the $\log(k_{ic})$ parameter for each subject within a condition is assumed to be a realization from a normal distribution with some population mean $\theta_c$ for its subjects, and variance $\frac{g\sigma^2}{T}$, i.e. the variance is constant across subjects, across conditions.

A call to `simulate_dataset` is shown below.

```{r}
set.seed(8678)
sim_data <- simulate_dataset(conditions = c("EFT", "HIT", "NCC"),
                             num_subj = c(120, 142, 172),
                             time_points = c(30, 90, 180, 365, 1095, 1825, 3650),
                             mean_ln_k = c(-6.877, -5.861, -6.078),
                             sigma_sq = 1.978, g = 10.407)
```

The `conditions` vector gives the name of the conditions in the simulated study.

The `num_subj` vector gives the sample size for each condition.

The `time_points` vector gives the surveyed delay lengths.

The `mean_ln_k` vector gives the condition mean hyperparameters.

The value `sigma_sq` gives the parameter $\sigma^2$, which represents the conditional variance of an observed transformed indifference point $\log(\frac{1}{\tilde D(t_{ijc})} - 1)$ given the subject's $\log(k_{ic})$ parameter.

The value `g` gives the ratio of the variance subject $\log(k_{ic})$ parameter to the conditional variance of the least squares (on the transformed scale) estimator of a subject's $\log(k_{ic})$ parameter given the true $\log(k_{ic})$. The result is that the true $\log(k_{ic})$ parameters have value $\frac{g\sigma^2}{T}$. (Recall $T$ is the length of the `time_points` vector.)

The data frame output by `sim_data` can be analyzed using our analysis methods. In particular, `prepare_data_frame` and `jb_rule_check` are ready to be run on the data frame. However, `dd_hyperbolic_model` and `get_subj_est_ln_k` will still need to be run on the output from `prepare_data_frame`.

```{r}
# These methods can be run on the output from simulate_dataset
prep_data <- prepare_data_frame(sim_data)
rule_check_results <- jb_rule_check(sim_data)

# These methods need to be run on the output from prepare_data_frame
subj_est_ln_k <- get_subj_est_ln_k(prep_data)
sim_model <- dd_hyperbolic_model(prep_data)
```

The output from `simulate_dataset` has one observation per simulated subject per delay for which an indifference point was observed. It has columns `subj`, `true_ln_k`, `group`, `delay`, and `indiff`.

The columns other than `true_ln_k` are necessary to run the analysis function on the dataset. The `subj` column identifies the subject, the `group` column identifies the condition, the `delay` column identifies the length of time for which the observed indifference point is measured, and the `indiff` column contains the subject's indifference point for the delay expressed as a proportion of the maximum reward.

The `true_ln_k` column contains the true $\log(k_{ic})$ parameter of the subject which the observation corresponds to. This can be used to compare estimates of the subject $\log(k_{ic})$ parameters with the true values.

For example, the following code produces a plot of the true vs estimated subject $\log(k_{ic})$ parameters, with points colored by rule check result.

```{r}
#| fig.alt: >
#|   Scatterplot showing the true vs estimated log k values for each subject,
#|   colored by whether or not the subject passed the rule check. The subject 
#|   log k parameter estimates are maximum likelihood estimates.
#|   The plot shows a line indicating where the estimated log k equals the true 
#|   log k, and the plotted points scatter around that line, since the estimator
#|   is unbiased. The true log k values of the rule check failures tend to vary
#|   more than the true log k values of the rule check passers.
# Get a data frame consisting of one observation per subject, containing each
#  subject's true ln(k)
subj_true_ln_k <- sim_data %>%
  select(subj, true_ln_k, group) %>%
  unique()

# Merge the data frames containing the true subject ln_k values, estimated
#  subject ln_k values, and rule check results.
subj_ln_k_comp <- merge(subj_est_ln_k, subj_true_ln_k)
subj_ln_k_comp_wrc <- merge(subj_ln_k_comp, rule_check_results)

# Get the overall rule check result for each subject
subj_ln_k_comp_wrc$rc_pass <- subj_ln_k_comp_wrc$C1 & subj_ln_k_comp_wrc$C2

# Make a plot showing true vs estimated ln(k) for each subject,
#  coloring the points based on whether or not the corresponding subject
#  passed the rule check.
plot(x = subj_ln_k_comp_wrc$true_ln_k, y = subj_ln_k_comp_wrc$ln_k,
     col = ifelse(subj_ln_k_comp_wrc$rc_pass, "green", "blue"),
     main = "Subject True vs Maximum Likelihood Estimated ln(k)",
     xlab = "True subject ln(k)", ylab = "Estimated subject ln(k)")
abline(a=0, b=1)
legend("topleft", 
       legend = c(
         "Passes rule check",
         "Fails rule check"),
       col = c("green", "blue"), pch = 1)
```

However, the data could be used to calculate other estimators of the subject $\log(k_{ic})$ for comparison purposes.

```{r}
#| fig.alt: >
#|   Scatterplot showing the true vs estimated log k values for each subject,
#|   colored by group. The log k estimates are a convex combination of the 
#|   subject parameter and condition hyperparameter maximum likelihood estimates
#|   chosen to minimize the mean square error of the subject log k estimate.
#|   The point positions are not visibly different since the estimates are
#|   nearly identical. The log k values in the E F T condition are slightly but
#|   noticably smaller on average compared to the H I T and N C C conditions. 
#|   The H I T and N C C conditions are not noticably different.

# Calculate alternative estimator for subj ln(k)
group_means <- sim_model$ln_k_mean %>%
  select(condition, ln_k_mean) %>%
  rename(group = condition)
g_hat <- sim_model$var["g"]

n_tp <- length(unique(sim_data$delay))

subj_ln_k_est_comp <- merge(subj_ln_k_comp, group_means) %>%
  rename(indiv_est_ln_k = ln_k,
         group_est_ln_k = ln_k_mean)

subj_ln_k_est_comp$reg_est_ln_k <- 
  (g_hat*subj_ln_k_est_comp$indiv_est_ln_k + 
     (1/n_tp)*subj_ln_k_est_comp$group_est_ln_k)/
  (g_hat + 1/n_tp)

# Plot alternative estimator vs true ln(k)
plot(x = subj_ln_k_est_comp$true_ln_k, y = subj_ln_k_est_comp$reg_est_ln_k,
     main = "Subject True vs MSE-Optimal Estimated ln(k)",
     col = ifelse(subj_ln_k_est_comp$group == "EFT", "darkgreen", 
                  ifelse(subj_ln_k_est_comp$group == "HIT", "orange", "skyblue")),
     xlab = "True subject ln(k)", ylab = "Estimated subject ln(k)")
abline(a=0, b=1)
legend("topleft", 
       legend = c(
         "EFT", "HIT", "NCC"),
       col = c("darkgreen", "orange", "skyblue"), pch = 1)
```
